---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# httr2

<!-- badges: start -->

[![R-CMD-check](https://github.com/r-lib/httr2/workflows/R-CMD-check/badge.svg)](https://github.com/r-lib/httr2/actions) [![Codecov test coverage](https://codecov.io/gh/r-lib/httr2/branch/master/graph/badge.svg)](https://codecov.io/gh/r-lib/httr2?branch=master)

<!-- badges: end -->

httr2 is a ground-up rewrite of [httr](https://httr.r-lib.org) that provides a pipeable API (with an explicit request object) and solves more problems that making API wrappers a pain (e.g. rate-limiting, retries, OAuth, storing secrets needed for testing, ...).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("r-lib/httr2")
```

## Usage

```{r}
library(httr2)
```

To use httr2, start by creating a request:

```{r}
req <- request("https://r-project.org")
req
```

You can tailor this request with the `req_` family of functions:

```{r}
# Add custom headers
req %>% req_headers("Accept" = "application/json")

# Add a body, turning it into a POST
req %>% req_body_json(list(x = 1, y = 2))

# Automatically retry if the request fails
req %>% req_retry(max_tries = 5)

# Change the HTTP method
req %>% req_method("PATCH")
```

You can see what httr2 will send to the server with `req_dry_run()`:

```{r}
req %>% req_dry_run(quiet = FALSE)
```

And perform the request, getting back a response, with `req_fetch()`:

```{r}
resp <- req_fetch(req)
resp
```

The `resp_` functions make it easy to work with the response:

```{r}
resp %>% resp_content_type()
resp %>% resp_status_desc()
resp %>% resp_body_html()
```

## Major differences to httr

-   HTTP errors are automatically converted into R errors.
    Use `req_error()` to override the defaults (which turn all 4xx and 5xx responses into errors) or to add additional details to the error message.

-   You can automatically retry if the request fails or encounters a transient HTTP error (e.g. a 429 rate limit request).
    `req_retry()` defines the maximum number of retries, which errors are transient, and how long to wait between tries.

-   OAuth support has been totally overhauled to directly support many more flows and to make it much easier to both customise the built-in flows and create your own.

-   You can manage encrypted secrets often needed for testing with `secret_encrpyt()` and friends.
    You can obfuscate mildly confidential data (like many client secrets) with `obfuscate()`, preventing them being scraped from published code.

## Acknowledgements

httr2 wouldn't be possible without [curl](https://jeroen.cran.dev/curl/), [openssl](https://github.com/jeroen/openssl/), [jsonlite](https://jeroen.cran.dev/jsonlite/), and [jose](https://github.com/jeroen/jose/), which are all maintained by [Jeroen Ooms](https://github.com/jeroen).
A big thanks also go to [Jenny Bryan](https://jennybryan.org) and [Craig Citro](https://research.google/people/CraigCitro/) who have given me much useful feedback on the both design of the internals and the user facing API.
