---
title: "httr2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{httr2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of this document is to get you up and running with httr2 as quickly as possible.
httr2 is designed to map closely to the underlying HTTP protocol.
I'll try and explain the basics in this intro, but I'd also recommend "[An overview of HTTP](https://developer.mozilla.org/en-US/docs/Web/HTTP/Overview)" from MDN.

There are two important parts to HTTP: the **request**, the data sent to the server, and the **response**, the data sent back from the server.

```{r setup}
library(httr2)
```

## Create a request

In httr2, you start by creating a request.
This is quite different to httr, where you could only submit a request, immediately receiving a response.
Having an explicit request object makes it easier to built up a complex request piece by piece, and works well with the pipe.

The simplest request just needs a url:

```{r}
req <- request("http://httpbin.org/get")
req
```

We can see exactly what this message this request will send to the server by performing a dry run:

```{r}
req %>% req_dry_run()
```

The first line of the request contains three important pieces of information:

-   The HTTP **method**, which a verb telling the server what action you want to perform.
    Here's its GET, indicating that we want to get some data.

-   The **path** which is the url stripped of all information that server already knows: the protocol (http), the domain or host (httpbin.org) and the port (not used here).

-   The version of the HTTP protocol.

The following lines consistent of name-value pairs separated by ":".
These are called the HTTP **headers**.
These headers are automatically added by httr2; if you want to add your own, you can use `req_headers()`:

```{r}
req %>%
  req_headers(Name = "Hadley", `Shoe-Size` = "11") %>% 
  req_dry_run()
```

HTTP servers will ignore headers that they don't understand.

The headers are finished up by a blank line which is followed by the **body**.
The requests above (like all GET requests) don't have a body, so let's see what happens if we add one.
Here we'll use `req_body_json()` to add some data encoded as JSON:

```{r}
req %>%
  req_body_json(list(x = 1, y = "a")) %>% 
  req_dry_run()
```

What's changed?

-   The method has changed from GET to POST.
    POST is the standard method for sending data to a website.
    You can use other methods with `req_method()`.

-   We have two new headers, `Content-Type` and `Content-Length` which tell the server how to interpret the body --- it's going to be 15 bytes long and is encoded as JSON.

-   We have a body, consisting of some JSON.

Different sites need data stored in different formats so httr2 provides a selection of common data types:

```{r}
req %>%
  req_body_form(list(x = 1, y = "a")) %>% 
  req_dry_run()
```

If you need to send data encoded differently, you can use `req_body_string()` or `req_body_raw()` to add the raw data to the body, and `req_header()` to set the `Content-Type` header to the appropriate type.

## Performing a request and fetch the response

To actually perform a request and fetch the response back, you'll use `req_fetch()`:

```{r}
req <- request("https://httpbin.org/json")
resp <- req %>% req_fetch()
resp
```

You can see a simulation of what httr2 actually received with `resp_raw()`:

```{r}
resp %>% resp_raw()
```

An HTTP response has a very similar structure to an HTTP request.
The first line gives the version of HTTP used, and a status code followed by a short description.
Then we have the headers, followed by a blank line, followed by a body.
The majority of responses will have a body, unlike requests.

You can extract data from the response using the `resp_()` functions:

-   `resp_status()` returns the status code and `resp_status_desc()` returns the description:

    ```{r}
    resp %>% resp_status()
    resp %>% resp_status_desc()
    ```

-   You can extract all headers with `resp_headers()` or a specific header with `resp_header()`:

    ```{r}
    resp %>% resp_headers() %>% str()
    resp %>% resp_header("Content-Length")
    ```

    Headers are case insensitive:

    ```{r}
    resp %>% resp_header("CoTtEnT-LeNgTH")
    ```

-   You can extract the body in various forms using the `resp_body_*()` family of functions.
    Since this response returns JSON we can use `resp_body_json()`:

    ```{r}
    resp %>% resp_body_json() %>% str()
    ```

Responses with status codes 4xx and 5xx are consider to be HTTP errors.
httr2 automatically turns these into R errors:

```{r, error = TRUE}
request("https://httpbin.org/status/404") %>% req_fetch()
request("https://httpbin.org/status/500") %>% req_fetch()
```

This is another important difference with httr, which required that you explicitly call `httr::stop_for_status()` to turn HTTP errors into R errors.
If needed, you can revert to the httr behaviour with `req_error(req, is_error = ~ FALSE)`.

## Controlling the request process

A number of `req_` functions don't directly affect how the request is sent but affect how it's handled.
I'm not going to go in to detail here, but wanted to make you aware of the options:

-   `req_cache()` sets up a cache so if repeated requests return the same results, you can avoid a trip to the server.

-   `req_throttle()` will automatically add a small delay before each request so you can avoid hammering a server with many requests.

-   `req_retry()` sets up a retry strategy so that if the request either fails or you get a transient HTTP error, it'll automatically retry after a short delay.
