---
title: "Wrapping APIs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Wrapping APIs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A common use for httr2 is wrapping up a useful API and exposing it in an R package where each API endpoint (i.e. a URL with parameters) becomes an R function with documented arguments.
This vignette will show you how, starting with a very simple API that doesn't need authentication, then slowly working up in complexity.

Along the way, you'll learn about how to:

-   Expose important details from HTTP errors in R errors.

-   Wrap up authentication details to provide a user friendly experience.

-   Consistently throttle the rate of requests or dynamically respond to rate limiting headers sent by the server.

I assume you're familiar with the basics of building a package.
If not, you might want to reading the "[The Whole Game](https://r-pkgs.org/whole-game.html)" chapter of [R packages](https://r-pkgs.org) first.

```{r setup}
library(httr2)
```

## Basics

```{r, include = FALSE}
# Seems to return 500s from time-to-time, so avoid any problems
# by only evaluating other chunks if a simple request succeeds.
faker_status <- request("https://fakerapi.it/api/v1") %>% 
  req_url_path_append("images") %>% 
  req_error(is_error = ~ FALSE) %>% 
  req_perform() %>% 
  resp_status()
faker_ok <- faker_status < 400
```

We'll start with a very simple API, [faker API](https://fakerapi.it/en){.uri}, which provides a collection of techniques for generating fake data.
Before we start writing the sort of functions that you might put in a package, we'll perform a request just to see how the basics work:

```{r, eval = faker_ok}
# We start by creating a request that uses the base API url
req <- request("https://fakerapi.it/api/v1")
resp <- req %>% 
  # Then we add on the images path
  req_url_path_append("images") %>% 
  # Add set _width and _quantity query parameters
  req_url_query(`_width` = 380, `_quantity` = 1) %>% 
  req_perform()

# The result comes back as JSON
resp %>% resp_body_json() %>% str()
```

### Errors

It's always worth a little early experimentation to see if we get any useful information from errors.
The httr2 details get in your way here because if you retrieve an unsuccessful HTTP response, you automatically get an error that prevents you from further inspecting the body:

```{r, error = TRUE, eval = faker_ok}
req %>% 
  req_url_path_append("invalid") %>% 
  req_perform()
```

To see what's going on, you need to start by using `req_error()` to suppress the automatic generation of R errors:

```{r, eval = faker_ok}
resp <- req %>% 
  req_url_path_append("imagesasf") %>% 
  req_error(is_error = ~ FALSE) %>% 
  req_perform()
```

Now you can look at the response body:

```{r, eval = faker_ok}
resp %>% resp_body_json()
```

It doesn't look like there's anything useful there so I'll double check the headers:

```{r, eval = faker_ok}
resp %>% resp_headers()
```

It doesn't look like we're getting any more useful information, so we can leave the `req_error()` default as is.
We'll have another go later with an API that does provide more details.

### User agent

When making a package that calls an API it's considered polite to set a user agent, so that if your app accidentally does something horribly wrong, the developers of the website can figure out who to reach out to.
You can do this with the `req_user_agent()` function:

```{r}
req %>%
  req_user_agent("httr2 (http://httr2.r-lib.org)") %>% 
  req_dry_run()
```

### Core request function

Once you've successful made a few requests, it's worth seeing if you can figure out the general pattern so you can wrap it up into a function that will become the core of your package.

For faker, I spent a little with the [documentation](https://fakerapi.it/en#basic-usage) noting some commonalities:

-   Every URL is of the form [`https://fakerapi.it/api/v1`](https://fakerapi.it/api/v1)`/{resource}`, and data is passed to the resource with query parameters.
    All parameters start with `_`.

-   Every resource has three common query parameters: `_locale`, `_quantity`, and `_seed` which have default values of `en_US`, `1` and `null` respectively.

-   All endpoints return JSON data.

This lead me to construct the following function:

```{r, eval = faker_ok}
faker <- function(resource, ..., quantity = 1, locale = "en_US", seed = NULL) {
  params <- list(
    ...,
    quantity = quantity,
    locale = locale,
    seed = seed
  )
  names(params) <- paste0("_", names(params))
  
  request("https://fakerapi.it/api/v1") %>% 
    req_url_path_append(resource) %>% 
    req_url_query(!!!params) %>% 
    req_user_agent("httr2 (http://httr2.r-lib.org)") %>% 
    req_perform() %>% 
    resp_body_json()
}

str(faker("images", width = 300))
```

There's one trick to notice here --- `req_url_query()` uses [dynamic dots](https://rlang.r-lib.org/reference/dyn-dots.html), so we can use `!!!` to convert (e.g.) `` req_url_query(req, !!!list(`_quantity` = 1, `_locale` = "en_US") `` into `` req_url_query(req, `_quantity` = 1, `_locale` = "en_US") ``.

Here I've decided to make my central function perform the request and parse the result.
If the API you are wrapping is more variable, you may want to return a request object that the caller can modify before fetching the response.

### Wrapping individual endpoints

`faker()` is quite general --- it's a good tool for the package developer because you can read the faker documentation and translate it to a function call.
But it's not very friendly for the package user who might not know anything about web APIs.
So typically the next step in the process is to wrap up some individual endpoints with their own functions.

For example, let's take the persons endpoint which has three additional parameters: `gender` (male or female), `birthday_start`, and `birthday_end`.
A simple wrapper would start something like this:

```{r, eval = faker_ok}
faker_person <- function(gender = NULL, birthday_start = NULL, birthday_end = NULL, quantity = 1, locale = "en_US", seed = NULL) {
  faker(
    "persons",
    gender = gender,
    birthday_start = birthday_start,
    birthday_end = birthday_end,
    quantity = quantity,
    locale = locale,
    seed = seed
  )  
}
str(faker_person("male"))
```

We could make it more user friendly by checking the input types, and returning the result as a tibble.
I did a quick and dirty conversion using purrr; depending on your needs you could use base R code or `tidyr::hoist()`.

```{r, eval = faker_ok}
library(purrr)

faker_person <- function(gender = NULL, birthday_start = NULL, birthday_end = NULL, quantity = 1, locale = "en_US", seed = NULL) {
  if (!is.null(gender)) {
    gender <- match.arg(gender, c("male", "female"))
  }
  if (!is.null(birthday_start)) {
    if (!inherits(birthday_start, "Date")) {
      stop("`birthday_start` must be a date")
    }
    birthday_start <- format(birthday_start, "%Y-%m-%d")
  }
  if (!is.null(birthday_end)) {
    if (!inherits(birthday_end, "Date")) {
      stop("`birthday_end` must be a date")
    }
    birthday_end <- format(birthday_end, "%Y-%m-%d")
  }
  
  json <- faker(
    "persons",
    gender = gender,
    birthday_start = birthday_start,
    birthday_end = birthday_end,
    quantity = quantity,
    locale = locale,
    seed = seed
  )  
  
  tibble::tibble(
    firstname = map_chr(json$data, "firstname"),
    lastname = map_chr(json$data, "lastname"),
    email = map_chr(json$data, "email"),
    gender = map_chr(json$data, "gender")
  )
}
faker_person("male", quantity = 5)
```

The next steps would be to export and document this function; but I'll leave that up to you.

## App with API key

```{r include = FALSE}
nyt_ok <- secret_has_key("HTTR2_KEY") && 
  # Don't run in R CMD check on GHA
  identical(Sys.getenv("RSPM"), "")
```

Next we'll take a look at the NYTimes [Books API](https://developer.nytimes.com/docs/books-product/1/overview). It requires a very simple authentication with an API key that's included in every request.
When you're wrapping an API that has a key you're going to face two struggles:

-   How do you test your package without sharing your key with the whole world?

-   How do you allow your users to supply their own key, without having to pass it to every function?

I'll start by tackling the first problem because otherwise there's no way for me to show how the API works in this vignette 😃.
We'll come back to the second at the very end of this section, because it's easiest to tackle once we have a function in place.

### Encrypting secrets

httr2 provides the `secret_decrypt()` and `secret_decrpyt()` functions that use asymmetric cryptography.
To use them you need to first create a "key" that's used to lock and unlock the secrets.
The best key is random because there's no way for anyone to guess it.
httr2 provides `secret_make_key()` to make a secure key for you:

```{r}
secret_make_key()
```

(Don't copy and paste this key; run `secret_make_key()` yourself to make your own personal key.)

Now you need to make this key available in the places were you want to unlock the secrets.
httr2 provides tools to retrieve keys from environment variables because this is a common way of communicating these sort of secrets.
So start by opening your `.Renviron` (e.g. with `usethis::edit_r_environ()`) and add a line like:

    YOURPACKAGE_KEY=key_that_you_generated_above

Another place you'll probably want to provide the key is in your Github actions.
This is a two step process:

1.  Add the key to your [repository secrets](https://docs.github.com/en/actions/reference/encrypted-secrets).

2.  Share the secret with the workflows that need it by adding a line to the appropriate workflow:

    ``` yaml
        env:
          HTTR2_KEY: ${{ secrets.HTTR2_KEY }}
    ```

    You can see how httr2 does it in <https://github.com/r-lib/httr2/blob/master/.github/workflows/R-CMD-check.yaml>.

Now you can encrypt your API key and store it in your tests --- there's no way for anyone to decrypt it unless they have your key.

```{r}
x <- secret_encrypt("my-secret", "HTTR2_KEY")
x
secret_decrypt(x, "HTTR2_KEY")
```

So now you can understand how the following code works:

```{r, eval = nyt_ok}
my_key <- secret_decrypt("4Nx84VPa83dMt3X6bv0fNBlLbv3U4D1kHM76YisKEfpCarBm1UHJHARwJHCFXQSV", "HTTR2_KEY")
```

Since most people won't have this key, n the source of this vignette I also check whether or not it's available with `nyt_ok <- secret_has_key("HTTR2_KEY")`, then use `eval = nyt_ok` in each chunk so that the code is only run when possible.

When `secret_decrypt()` is run in tests, it will automatically skip the test if the key isn't found.
That means your tests will automatically pass on CRAN.

### Basic request

Now lets perform a test request and look at the response:

```{r, eval = nyt_ok}
resp <- request("https://api.nytimes.com/svc/books/v3") %>% 
  req_url_path_append("/reviews.json") %>% 
  req_url_query(`api-key` = my_key, isbn = 9780307476463) %>% 
  req_perform()
resp
```

Like most modern APIs, this one returns the results as JSON:

```{r, eval = nyt_ok}
resp %>% 
  resp_body_json() %>% 
  str()
```

Before we start wrapping this up into a function lets consider the security implications and what happens with errors.

### Security considerations

Note that including an API key in a query is a relatively insecure method; if an API uses it its typically because the key is relatively easy to create or gives relatively few privileges.
Here it only takes a couple of minutes to generate your own NYTimes API key, so there's little incentive for someone to try and steal yours.

Section 5.3 of [rfc6750](https://datatracker.ietf.org/doc/html/rfc6750#section-5.3) has this to say about passing bearer tokens[^1] in the URL:

[^1]: A bearer token is a little different to an access key but the distinction isn't important here.

> **Don't pass bearer tokens in page URLs**: Bearer tokens SHOULD NOT be passed in page URLs (for example, as query string parameters).
> Instead, bearer tokens SHOULD be passed in HTTP message headers or message bodies for which confidentiality measures are taken.
> Browsers, web servers, and other software may not adequately secure URLs in the browser history, web server logs, and other data structures.
> If bearer tokens are passed in page URLs, attackers might be able to steal them from the history data, logs, or other unsecured locations.

httr2 makes no efforts to redact confidential information stored in query parameters.
This means it's relatively easy to leak your key if you use `req_perform(verbose = 1)` or `req_dry_run()`.
This just means you'll need to be a little cautious if you need to get help with a reprex; you'll need to manually redact the key.

### Error handling

What happens if there's an error?
For example, if we deliberately supply an invalid key:

```{r, eval = nyt_ok, error = TRUE}
resp <- request("https://api.nytimes.com/svc/books/v3") %>% 
  req_url_path_append("/reviews.json") %>% 
  req_url_query(`api-key` = "invalid", isbn = 9780307476463) %>% 
  req_perform()
```

To see if there's any extra useful information we can again suppress the default error and inspect the result:

```{r, eval = nyt_ok, error = TRUE}
resp <- request("https://api.nytimes.com/svc/books/v3") %>% 
  req_url_path_append("/reviews.json") %>% 
  req_url_query(`api-key` = "invalid", isbn = 9780307476463) %>% 
  req_error(is_error = ~ FALSE) %>% 
  req_perform()
resp
resp %>% resp_body_json()
```

It looks like there's some useful additional info in the `faultstring`:

```{r, eval = nyt_ok}
resp %>% resp_body_json() %>% .$fault %>% .$faultstring
```

To add that information to future errors we can use the `info` argument to `req_error()`.
Once we do that and re-fetch the request, we see the additional information displayed in the R error:

```{r, eval = nyt_ok, error = TRUE}
resp <- request("https://api.nytimes.com/svc/books/v3") %>% 
  req_url_path_append("/reviews.json") %>% 
  req_url_query(`api-key` = "invalid", isbn = 9780307476463) %>% 
  req_error(info = function(resp) {
    resp %>% resp_body_json() %>% .$fault %>% .$faultstring
  }) %>% 
  req_perform()
```

### Rate limits

The [frequently asked questions](https://developer.nytimes.com/faq#a11) also mentions rate limits:

> Yes, there are two rate limits per API: 4,000 requests per day and 10 requests per minute.
> You should sleep 6 seconds between calls to avoid hitting the per minute rate limit.
> If you need a higher rate limit, please contact us at [code\@nytimes.com](mailto:code@nytimes.com){.email}.

Since the docs don't mention that the use of `Retry-After` header, we'll use `req_throttle()` to ensure we don't make more than 10 requests every 60s:

```{r, eval = nyt_ok}
req <- request("https://api.nytimes.com/svc/books/v3") %>% 
  req_url_path_append("/reviews.json") %>% 
  req_url_query(`api-key` = "invalid", isbn = 9780307476463) %>% 
  req_throttle(10 / 60)
```

This limit is automatically shared across all requests made to `api.nytimes.com`.
Since the docs suggest the rate limit applies per API, you might want to use the `realm` argument to be a bit more specific:

```{r, eval = nyt_ok}
req <- request("https://api.nytimes.com/svc/books/v3") %>% 
  req_url_path_append("/reviews.json") %>% 
  req_url_query(`api-key` = "invalid", isbn = 9780307476463) %>% 
  req_throttle(10 / 60, realm = "https://api.nytimes.com/svc/books")
```

Many APIs return additional information about how long to wait when the rate limit is exceeded.
So I deliberately violated the rate limit by quickly making 11 requests; unfortunately while the response was a standard 429 (Too many requests), it did not include any information about how long to wait in either the response body or the headers.

### Wrapping it up

Putting together all the pieces above yields a function something like this:

```{r, eval = nyt_ok}
nytimes_books <- function(api_key, path, ...) {
  request("https://api.nytimes.com/svc/books/v3") %>% 
    req_url_path_append("/reviews.json") %>% 
    req_url_query(..., `api-key` = api_key) %>% 
    req_error(info = function(resp) {
      resp %>% resp_body_json() %>% .$fault %>% .$faultstring
    }) %>% 
    req_throttle(10 / 60, realm = "https://api.nytimes.com/svc/books") %>% 
    req_perform() %>% 
    resp_body_json()
}

drunk <- nytimes_books(my_key, "/reviews.json", isbn = "0316453382")
drunk$results[[1]]$summary
```

Again, for a real package, you'd want to spend some time converting the nested list into a more user friendly data structure.

### API key

If you were going to put this into a package, you'd need to think about some convenient interface for supplying the API key.
An environment variable is a good place to start because they're easy to set without typing anything in the console (which can get accidentally shared via your `.Rhistory`) and are easily set in automated processes.

```{r, eval = nyt_ok}
get_api_key <- function() {
  key <- Sys.getenv("NYTIMES_KEY")
  if (identical(key, "")) {
    stop("No API key found, please supply with `api_key` argument or with NYTIMES_KEY env var")
  }
  key
}

req_nytimes_books <- function(path, ..., api_key = get_api_key()) {
  ...
}
```

Since the `api_key` is no longer required, we can make it the last argument to the function.
It's still available to override and document, but it'll only be needed in exceptional circumstances.

## Sending data & HTTP methods

<https://docs.github.com/en/rest/reference/gists#star-a-gist>

Just use basic auth: <https://docs.github.com/en/rest/overview/other-authentication-methods#basic-authentication> username + token.
Good example of why tokens: you can have many different tokens, each which can do different things, and can be revoked without affecting any other token or requiring you to change your password.

Need to be a little careful with tokens in env vars.
Any R package could use.
But this doesn't really pose any additional threat: an R package can run any code, so it's already very powerful.
If you are worried about this attack vector you need to resolve it by ensuring that the process that runs R is locked down (e.g. run in a container or other VM).

Rate limiting: <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>

## App that uses OAuth

If the API provides access to a website where the user already has an account (think twitter, instagram, facebook, google, etc), it's likely to use OAuth to allow you to authenticate on behalf of the user.
OAuth[^2] is an authentication framework that's designed so that you don't have to share your username and password with an app; instead the app asks for permission to use your account.
You've almost certainly used this before on the web; it's used in most cases where one website wants to use another website on your behalf.

[^2]: Here I'll only talk about OAuth 2.0 which is the only version in common use today.
    OAuth 1.0 is largely only of historical interest.

### OAuth clients

The first step in working with any OAuth api is to create a client or app.
This involves you registering for a developer account on the APIs website and creating a new OAuth app.
The process varies from API to API, but at the end of it you'll be a client id and in most cases a client secret.

If the API provides a way to authenticate without the client secret, you should leave it out of your package.
But in most cases, you'll need to include the secret in the package.
You can use `obfuscate()` to hide the secret; again this is not bullet proof but for most people it will be easier to create a new client than try and reuse yours.
It is unusual for an OAuth client to be able to do anything in its own right, so even if someone does steal your secret there's not much harm they can do with it (generally the worst they can do is use up your rate limits).

To obfuscate a secret, call `obfuscate()`:

```{r}
obfuscate("secret")
```

Then use the client id from the website along with the obfuscated secret to create a client:

```{r}
client <- oauth_client(
  id = "client_id", 
  secret = obfuscated("YVW06Lex"),
  token = "https://accounts.google.com/o/oauth2/token"
)
client
```

Note that when you print the client, httr2 automatically redacts the secret --- this is part of the general effort that httr2 makes to prevent you from accidentally exposing any confidential information in logs or reprexes.

### OAuth app

Once you have a client, you'll need to collect a little more information to create an OAuth app.
An app is the combination of the information about your client and some general information about the API.
(From the perspective of the API client and app are synonymous because it's only thinking of itself).

The information you'll need varies a little depending on the flow you're using, but for the most common case it will look something like this.
You'll pass the client that you created above, as long as a list of authentication endpoints.
In the most common case, you'll need to supply `token` and `authorization` endpoints.

Depending on how the API is set up, you may also need to change the `auth` parameter.
We'll come back to that shortly.

### OAuth flows

OAuth provides a number of different "flows".
The most common is the "authorisation code" flow, which is implemented by `req_oauth_auth_code()`.
It goes something like this:

-   Client makes authorisation request to resource owner, and it returns an authorisation grant (four types are defined in this rfc: authorisation code, implicit, resource owner password credentials, and client credentials).

-   Client makes authorisation grant to authorisation server (which can be different to resource server, even though that appears to be rare in practice) which returns access token.

-   Client uses access token to access protected resources.
    `req_auth_bearer_token()`

Other flows provide different ways getting the token:

-   `req_oauth_client_credentials()` is used to allow the client to perform actions on its own behalf (instead of on behalf of some other user).

-   `req_oauth_device()` uses the "device" flow which is designed for devices like TVs that don't have an easy way to enter data.
    It also works well from the console.

-   `req_oauth_bearer_jwt()` uses a JWT signed by a private key.

-   `req_oauth_password()` exchanges a user name and password for an access token.

-   `req_oauth_refresh()` works directly with a refresh token that you've got elsewhere.
    It's useful for testing.

There's one historically important OAuth flows that httr2 doesn't support: the implicit grant flow.
This is now [mostly deprecated](https://developer.okta.com/blog/2019/05/01/is-the-oauth-implicit-flow-dead) and was never a particularly good fit for native applications because it relies on a technique for returning the access token that only works inside a web browser.

When wrapping an API, you'll need to carefully read the documentation to figure out which flows are available.
Typically you'll want to use the auth code flow, but if it's not available you'll need to carefully consider the others.

An additional wrinkle is that many APIs don't implement the flow in exactly the same way as the spec.
If your initial attempt doesn't work, you're going to need to do some sleuthing.
This is going to be moderately painful but unfortunately there's no way around it.
I recommend using `with_verbose()` so you can see exactly what httr2 is sending to the server.
You'll then need to carefully compare this to the API documentation and play spot the difference.

To get started, I recommend first working with the lower-level `oauth_flow_*()` functions.
They just return a token (rather than authenticating a request) so they're easy to work with as you get the system up and running.

```{r, eval = FALSE}
```

### Token caching

Token caching.
Many access tokens also have a refresh token, which is a way to get new access tokens.
httr2 can cache these tokens.
If the token has expired and has a refresh token, we automatically refresh before the first request.
If the request errors with a 401 and error `invalid_token` then we try the refresh token if available, otherwise re-run the full flow.
