---
title: "OAuth"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(httr2)
```

OAuth[^1] is an authorization framework design for performing work on the behalf of a user.
It allows you to wrap an API that requires a user login without your app ever directly seeing the user's user name and password.
You've almost certainly used this before on the web or with your TV; it's used in most cases where one website (or your TV) needs to login to an account on your behalf.

[^1]: Here I'll only talk about OAuth 2.0 which is the only version in common use today.
    OAuth 1.0 is largely only of historical interest.

OAuth is a broad framework that has many different variants, called **flows**, which makes it hard to provide generalisable advice.
The one commonality is that all flows eventually generate an **access token** (essentially a random string of numbers and letters) which is provided to the API in the form of a **bearer token** in the `Authorization` header.
If you have this access token, you can use it directly with `req_auth_bearer_token()`, but in most cases you will want to give httr2 the information to generate and use the access token for you by calling a `req_oauth_` function.
The rest of the vignette describes just how to do that.

## Clients

The first step in working with any OAuth API is to create an **application** **client**, which I'll generally just call a **client**.
You'll do this by registering for a developer account on the API's website then creating a new OAuth app.
This process varies from API to API (it's normal to spend some time hunting through the docs and settings), but at the end of it you'll get a client id (another random string of numbers and letters).

Sometimes the client id is all you need, and you can give httr2 the information about the client with a call to `oauth_client()`, something like this:

```{r}
client <- oauth_client(
  id = "28acfec0674bb3da9f38",
  token_url = "https://github.com/login/oauth/access_token",
  name = "hadley-oauth-test-2"
)
```

The call to `oauth_client()` also includes a `name` and a `token_url`.
The `name` is human-facing, and should typically be the same as your package or project (the thing that prompted you to create the client).
I have a bunch of apps that I've used for testing, so here I've and is used to remind you which app the client corresponds to.
The `token_url` points to the URL that's used to obtain an access token.
You'll need to find this from the documentation for the API you're wrapping; it will typically be found in the section that describes the OAuth process and will be an endpoint that returns an access token.

### Client secret

In most cases, however, the API will also require a client secret.
While this is called a secret, it's typically not that important to keep it a secret because of two reasons:

-   It's typically easy to create an new app on the developer website so stealing yours wouldn't save much time.
-   It's unusual for an OAuth client to be able to do anything in its own right, so stealing your secret doesn't have much benefit.

That said, httr2 provides some tooling to obfuscate your client secret so that the client secret isn't directly embedded in your source code, and hence vulnerable to scraping off GitHub or similar.

To obfuscate a string, call `obfuscate()` then copy and paste the result into your package.
For example, if your client secret was "secret", you'd call `obfuscate()` then you'd copy and paste `obfuscated("B4Evdd5x4wl0XTWvtTpuGaw7nM7GEg")` into your client specification.

```{r}
obfuscate("secret")
```

Here's what a complete client specification for GitHub looks like, using a real app that I created specifically for this vignette:

```{r}
client <- oauth_client(
  id = "28acfec0674bb3da9f38",
  secret = obfuscated("J9iiGmyelHltyxqrHXW41ZZPZamyUNxSX1_uKnvPeinhhxET_7FfUs2X0LLKotXY2bpgOMoHRCo"),
  token_url = "https://github.com/login/oauth/access_token",
  name = "hadley-oauth-test"
)
```

You can certainly uncover my client secret if you are an experienced R programmer and are willing to spend a bit of time experimenting, but I'm pretty sure it'd be easy for you to just create your own app on GitHub.

### Packaging

I recommend wrapping client creation in a function in your package, e.g.:

```{r}
github_client <- function() {
  oauth_client(
    id = "28acfec0674bb3da9f38",
    secret = obfuscated("J9iiGmyelHltyxqrHXW41ZZPZamyUNxSX1_uKnvPeinhhxET_7FfUs2X0LLKotXY2bpgOMoHRCo"),
    token_url = "https://github.com/login/oauth/access_token",
    name = "hadley-oauth-test"
  )
}
```

You'll need this in order to run tests for your package, but you'll probably also want to use it as the default client for your users.
In some cases, it will be necessary for each user to create their own app and matching client (e.g. if rate limits are applied by app, not by user), but this is much less user friendly so you should avoid it if possible.
That said, you should always provide a way for a user to supply their own client when you bundle a default.

## Authorization code flow

Once you have a client you need to use it with a flow in order to get a token.
You'll need to read the docs for your API to figure out which flows it supports, but the most common is the **authorization code**[^2] flow, which works something like this:

[^2]: Confusingly, Github calls this the "web application flow", but this is not the official name.
    This is a general problem with OAuth docs; often they use different terms for the same things.
    In httr2, I've tried to stick as closely as possible to the terms used in the official RFC specifications.

1.  httr2 sends the user to an **authorization URL** provided by the API. The URL includes parameters that identify your app and what **scope** of access you're looking for.
2.  The user signs in and approves the request.
3.  The API sends an **authorization code** back to httr2 using a callback URL that was supplied in the initial request.
4.  httr2 sends the authorization code to the **token URL** to get an access token.

In httr2 this flow is implemented by a pair of functions: `oauth_flow_auth_code()` and `req_oauth_auth_code()`.
Start with `oauth_flow_auth_code()` to check that you have all the parameters correctly specified, then use `req_oauth_auth_code()` to authenticate requests.
These two steps are described in the following sections.

### Creating a token

``` oauth_flow_auth_code(``) ``` is the best way to verify you've correctly specified the parameters to the client and the `auth_url`, without depending on correclty understanding any other part of the API.
For example, you could get a token to access the GitHub API (using the client defined above) with this code:

```{r, eval = FALSE}
token <- oauth_flow_auth_code(
  client = client, 
  auth_url = "https://github.com/login/oauth/authorize"
)
```

This flow can't be used inside a vignette because it's designed specifically for interactive use, but if you do run it and print out the `token`, you'll see something like this:

```{r}
#| eval: false
token
#> <httr2_token>
#> token_type: bearer
#> access_token: <REDACTED>
#> scope: ''
```

There's not much to see here because httr2 automatically redacts the access token (because that could be used to perform actions on behalf of the user).

If your call to `oauth_flow_auth_code()` succeeds then you've got everything set up correctly and you can proceed to the next step.
Otherwise, you'll get an HTTP error.
If you're very lucky, that error will be informative and will help you figure out want went wrong.
However, in most cases, you'll need to carefully double check that you've correctly copied and pasted the client id and secret, and check that you've supplied the correct authorization and token urls (`auth_url` and `token_url`).
If the docs have multiple candidates for each and you're unclear about which is which, you'll need to do some systematic experimentation.

### Authenticating a request

Initial configuration is the only time that you'll see an `httr2_token` object because you'll generally want to rely on httr2 to manage the tokens for you.
You'll do that with `req_oauth_auth_code()`. 
To check that it's working correctly, I recommend finding the simplest possible API endpoint to test it with.
A good place to start is an endpoint that provides information about the "current" user, if your API provides one.

For example, the GitHub API provides a `GET` endpoint at `/user` that returns information about the current user.
If we make a request to this endpoint without authentication, we'll get an error:

```{r}
#| error: true
req <- request("https://api.github.com/user")
req %>% 
  req_perform()
```

We can authenticate this request with `req_oauth_auth_code()`, using the same arguments as our previous call to `oauth_flow_auth_code()`:

```{r}
#| eval: false
req %>% 
  req_oauth_auth_code(
    client = github_client(),
    auth_url = "https://github.com/login/oauth/authorize"
  ) %>% 
  req_perform() %>% 
  resp_body_json() %>% 
  str()
```

When you run this code, you'll see something like this, but it will obviously contain information about you, not me.

```{r}
#| eval: false
#> List of 32
#>  $ login        : chr "hadley"
#>  $ id           : int 4196
#>  $ node_id      : chr "MDQ6VXNlcjQxOTY="
#>  $ avatar_url   : chr "https://avatars.githubusercontent.com/u/4196?v=4"
#>  $ gravatar_id  : chr ""
#>  $ url          : chr "https://api.github.com/users/hadley"
#>  $ html_url     : chr "https://github.com/hadley"
#>  ...
#>  $ type         : chr "User"
#>  $ site_admin   : logi FALSE
#>  $ name         : chr "Hadley Wickham"
#>  $ company      : chr "@posit-pbc"
#>  $ blog         : chr "http://hadley.nz"
#>  $ location     : chr "Houston, TX"
```

### Token management

There are two big reasons to allow httr2 to manage tokens for you.
The first is that httr2 will automatically refresh the token if it's expired.
The second is cross-session caching which you can turn on with `cache_disk = TRUE`.
Turning it on means that your users will only need to re-authenticate once a month, but does means their credentials are saved on disk.
Reading the "Caching" section in `?req_oauth_auth_code` for more details.

## Other flows

When wrapping an API, you'll need to carefully read the documentation to figure out which flows it provides.
If possible you'll want to use the authorization code flow since it generally provides the best experience, but if it's not available you'll need to carefully consider the others.
Currently, httr2 supports the following flows:

-   `req_oauth_device()` uses the "device" flow which is designed for devices like TVs that don't have an easy way to enter data.
    It also works well from the console.

-   `req_oauth_client_credentials()` and `req_oauth_bearer_jwt()` are often needed for **service accounts**, accounts that represent automated services, not people, and are often used in non-interactive environments.

-   `req_oauth_password()` exchanges a user name and password for an access token.

-   `req_oauth_refresh()` works directly with a refresh token that you already have.
    It's useful for testing and automation.

httr2 doesn't support the implicit grant flow.
This was historically important but is now [mostly deprecated](https://developer.okta.com/blog/2019/05/01/is-the-oauth-implicit-flow-dead) and was never a particularly good fit for R because it relies on a technique for returning the access token that only reliably works inside a web browser.

Regardless of which flow you use, you'll need to follow the same process as the example above: first figure out how to get a token using the `oauth_flow_` function, then actually use oauth with a request, by calling the matching `req_oauth_` function.

One additional wrinkle is that many APIs don't implement the flow in exactly the same way as the spec so httr2's built-in flows might not work at all.
If your initial attempt doesn't work, you're going to need to do some sleuthing.
This is going to be painful, but unfortunately there's no way around it.
I recommend wrapping your httr2 code in `with_verbosity()` so you can see exactly what httr2 is sending to the server.
You'll then need to carefully compare this to the API documentation and play "spot the difference".
You're very welcome to file an issue and I'll do my best to help you out.
